Price of Anarchy for Mean Field Games

### 2018-08-02 PoA.tex
stylistic edit for section 2.1
rearranging the position of Theorem 2, make it before the new notation of xt_bar and vt_bar

\subsection{\textbf{Linear Quadratic Extended Mean Field Games}}\label{sec:EMFG}
To solve the linear quadratic extended mean field game (LQEMFG), we begin by considering the reduced Hamiltonian for this problem:
\begin{equation*}
\begin{split}
    H(t,x,\bar{\mu},\alpha,\bar{\nu},y)=&\left[b_1(t)x+\bar{b}_1(t)\bar{\mu}+b_2(t) \alpha+\bar{b}_2(t)\bar{\nu}\right]y \\
    &+\frac{1}{2}\left[q(t)x^2+\bar{q}(t)(x-s(t)\bar{\mu})^2 +r(t)\alpha^2+\bar{r}(t)(\alpha-\bar{s}(t)\bar{\nu})^2\right],
\end{split}
\end{equation*}
and whenever the flows $\bmubar=(\bar{\mu}_t)_{0\leq t\leq T}$ and $\bnubar=(\bar{\nu}_t)_{0\leq t\leq T}$ are fixed, we consider for each control process $\balpha=(\alpha_t)_{0\leq t\leq T}$ the adjoint equation:
\begin{equation*}
    dY_t=- \partial_x H(t,X_t,\bar{\mu}_t,\alpha_t,\bar{\nu}_t,Y_t)dt+Z_tdW_t,
    \qquad
    Y_T=\partial_xg(X_T, \cL(X_T)).
\end{equation*}
According to the Pontryagin stochastic maximum principle, a sufficient condition for optimality is $\partial_{\alpha}H(t,X_t,\bar{\mu}_t,\hat{\alpha}_t,\bar{\nu}_t,y)=0$. We introduce the function:
\begin{equation}
    \hat{\alpha}(t,x,\bar{\mu},\bar{\nu},y)=\frac{\bar{r}(t)\bar{s}(t)\bar{\nu}-b_2(t)y}{r(t)+\bar{r}(t)},
\label{eq:MFG_opt_alpha_Pontryagin}
\end{equation}
and use the control $\hat{\alpha}_t=\hat{\alpha}(t,X_t,\bar{\mu},\bar{\nu},Y_t)$. When solving the fixed point step, we identify $\bar{\nu}_t=\mathbb{E}(\hat{\alpha}_t)$. By taking the expectation, we find:
\begin{equation*}
    \bar{\nu}_t = \mathbb{E}(\hat{\alpha}_t)=c^{MFG}(t)\mathbb{E}(Y_t),
	\qquad 
	\text{with} 
	\quad
    c^{MFG}(t)=-\frac{b_2(t)}{r(t)+\bar{r}(t)(1-\bar{s}(t))}.
\end{equation*}
Thus, from equation (\ref{eq:MFG_opt_alpha_Pontryagin}) we must have:
\begin{equation}
    \hat{\alpha}_t=a^{MFG}(t) Y_t+b^{MFG}(t)\mathbb{E}(Y_t),
\label{eq:MFG_opt_alpha_Y_expY}
\end{equation}
where
\begin{equation*}
	a^{MFG}(t)=-\frac{b_2(t)}{r(t)+\bar{r}(t)}
	\qquad
	\text{and}
	\qquad
	b^{MFG}(t)=-\frac{\bar{r}(t)\bar{s}(t)b_2(t)}{(r(t)+\bar{r}(t))(r(t)+\bar{r}(t)(1-\bar{s}(t)))}.
\end{equation*}
Note that $c^{MFG}(t)=a^{MFG}(t)+b^{MFG}(t)$. The solution of the mean field game equilibrium problem is given by the solution to the FBSDE system:
\begin{equation}
\left\{
\begin{array}{lcl}
      \displaystyle  dX_t&=& \left[b_1(t)X_t+\bar{b}_1(t) \mathbb{E}X_t+a^{MFG}(t)b_2(t)Y_t+(b^{MFG}(t)b_2(t)+c^{MFG}(t)\bar{b}_2(t))\mathbb{E}Y_t\right]dt+\sigma dW_t \\[5pt]
        dY_t&=& -\left[(q(t)+\bar{q}(t))X_t-\bar{q}(t)s(t)\mathbb{E}X_t+b_1(t)Y_t \right]dt +Z_t dW_t,
\end{array}
\right.
\label{eq:FBSDE_EMFG}
\end{equation}
with initial condition $X_0=\xi$, a random variable with finite mean and variance, and terminal condition $Y_T=(q_T+\bar{q}_T)X_T-\bar{q}_Ts_T\mathbb{E}X_T$.

@@

We study in the following the variation of $PoA$ by letting one of the coefficients tend to zero or to infinity. It will be useful for us to note here the scalar Riccati equations associated with $u_t := \lambda \bar{\eta}_t^{MFG}$, $w_t := \bar{\eta}_t^{MKV}$ and $\eta_t$:
    \begin{equation}
		\dot{u}_t - 2A^u u_t - B u_t^2 + C^u =0, \qquad u_T = D^u,
		\label{eq:riccati_ut}
	\end{equation}	
	\begin{equation}
	    \dot{w}_t - 2A^w w_t - B w_t^2 + C^w =0, \qquad w_T = D^w,
	\label{eq:riccati_wt}
	\end{equation}
	\begin{equation}
	    \dot{\eta}_t - 2A^\eta \eta_t - B^\eta \eta_t^2 + C^\eta =0, \qquad \eta_T = D^\eta,
	\label{eq:riccati_etat}
	\end{equation}
	with:
	\begin{IEEEeqnarray*}{rClcrClcrCl}
		A^u &=& - \left(b_1 + \frac{\bar{b}_1}{2}\right), & \qquad & 
		A^w &=& - (b_1 + \bar{b}_1), & \qquad & 
		A^\eta &=& - b_1, \nonumber \\
		B^u &=& \frac{(b_2+\bar{b}_2)^2}{r + \bar{r}(1-\bar{s})^2}, & \qquad &	
		B^w &=& \frac{(b_2+\bar{b}_2)^2}{r + \bar{r}(1-\bar{s})^2}, &\qquad & 
		B^\eta &=& \frac{b_2^2}{r+ \bar{r}}, \nonumber \\
		C^u &=& \lambda (q + \bar{q}(1-s)),& \qquad & 
		C^w &=& q + \bar{q}(1-s)^2, & \qquad &
		C^\eta &=& q + \bar{q}, \nonumber \\
		D^u &=& \lambda(q_T + \bar{q}_T(1-s_T)), & \qquad & 
		D^w &=& q_T + \bar{q}_T(1-s_T)^2, & \qquad & 
		D^\eta &=& q_T + \bar{q}_T .
	\end{IEEEeqnarray*}
If $B^u\neq 0$, $B^u D^u \geq 0$ and $B^uC^u >0$, we have (see equation (\ref{eq:riccati_ut_sol_app}) in Appendix A) the existence and uniqueness for $u_t$ which can be expressed by:
\begin{equation}
		u_t= \frac{C^u(1-e^{-(\delta_u^+ - \delta_u^-)(T-t)})+D^u(\delta^+_u -\delta^-_ue^{-(\delta_u^+ - \delta_u^-)(T-t)} )}{B^uD^u(1-e^{-(\delta_u^+ - \delta_u^-)(T-t)})+ \delta^+_ue^{-(\delta_u^+ - \delta_u^-)(T-t)} -\delta^-_u  },
	\label{eq:riccati_ut_sol}
\end{equation}
with $\delta^\pm_u = -A^u \pm \sqrt{(A^u)^2 + B^u C^u}$. Under assumption (\ref{eq:assumptions}), the above conditions on $B^u$, $C^u$, and $D^u$ are satisfied, and we have $\delta^-_u<0<\delta^+_u$, $u_t > 0$ for all $t\in[0,T)$, and $u_T \geq 0$. We have analogous expressions for $w_t$ and $\eta_t$, in terms of $\delta^\pm_w$ and $\delta^\pm_{\eta}$, respectively. Note that $B^u=B^w=:B$.

To make the following computations easier to follow, we repeat equations (\ref{eq:PoA_LQ}), (\ref{eq:x_bar_MFG_explicit}), and (\ref{eq:v_t}), which we recall is equivalent to  equation (\ref{eq:v_t_MKV}), using the above notations:
\begin{equation}
PoA =1+ \frac{\Delta SC}{SC^{MKV}}= 1 +  \frac{ \displaystyle B\int_0^T (u_t-w_t)^2 \cdot (\bar{x}_t^{MFG})^2 dt }{\displaystyle \int_0^T \left[q+\bar{q} + B^{\eta} \eta_t^2 \right] v_t dt + (q_T + \bar{q}_T) v_T + w_0 (\mathbb{E}(\xi))^2},
\label{eq:PoA_LQ_new_notation}
\end{equation}
\begin{equation}
\bar{x}^{MFG}_t=\mathbb{E}(\xi) e^{\int_0^t(b_1+\bar{b}_1-Bu_s)ds},
\label{eq:x_bar_MFG_explicit_new_notation}
\end{equation}
\begin{equation}
v_t=Var(\xi)e^{\int_0^t 2(b_1-B^{\eta} \eta_s)ds}+\sigma^2 \int_0^t e^{2 \int_s^t (b_1-B^{\eta}\eta_u) du}ds.
\label{eq:v_t_new_notation}
\end{equation}
Also for convenience, recall the definition from equation (\ref{eq:lambda}):
\begin{equation*}
\lambda = \frac{b_2}{b_2 + \bar{b}_2}\cdot \frac{ r + \bar{r}(1-\bar{s})^2 }{r + \bar{r}(1-\bar{s})}.
\end{equation*}

@@

In the following propositions, we utilize the following assumption.	
\begin{assumption}\label{assumption}
	Assume (\ref{eq:assumptions}). In addition, assume: $b_1 > 0$, $D^{u} > 0$, $D^w > 0$, $D^{\eta} >0$ and the initial condition satisfies $\mathbb{E}(\xi) \neq 0$.
\end{assumption}

>>>>

\subsection{\textbf{Linear Quadratic Extended Mean Field Games}}\label{sec:EMFG}
To solve the linear quadratic extended mean field game (LQEMFG), we begin by considering the reduced Hamiltonian for this problem:
\begin{equation*}
\begin{split}
    H(t,x,\bar{\mu},\alpha,\bar{\nu},y)=&\left[b_1(t)x+\bar{b}_1(t)\bar{\mu}+b_2(t) \alpha+\bar{b}_2(t)\bar{\nu}\right]y \\
    &+\frac{1}{2}\left[q(t)x^2+\bar{q}(t)(x-s(t)\bar{\mu})^2 +r(t)\alpha^2+\bar{r}(t)(\alpha-\bar{s}(t)\bar{\nu})^2\right],
\end{split}
\end{equation*}
and whenever the flows $\bmubar=(\bar{\mu}_t)_{0\leq t\leq T}$ and $\bnubar=(\bar{\nu}_t)_{0\leq t\leq T}$ are fixed, we consider for each control process $\balpha=(\alpha_t)_{0\leq t\leq T}$ the adjoint equation:
\begin{equation*}
    dY_t=- \partial_x H(t,X_t,\bar{\mu}_t,\alpha_t,\bar{\nu}_t,Y_t)dt+Z_tdW_t,
    \qquad
    Y_T=\partial_xg(X_T, \cL(X_T)).
\end{equation*}
According to the Pontryagin stochastic maximum principle, a sufficient condition for optimality is $\partial_{\alpha}H(t,X_t,\bar{\mu}_t,\hat{\alpha}_t,\bar{\nu}_t,y)=0$. By solving this equation, we introduce the function:
\begin{equation}
    \hat{\alpha}(t,x,\bar{\mu},\bar{\nu},y)=\frac{\bar{r}(t)\bar{s}(t)\bar{\nu}-b_2(t)y}{r(t)+\bar{r}(t)},
\label{eq:MFG_opt_alpha_Pontryagin}
\end{equation}
and pick an optimal control as $\hat{\alpha}_t=\hat{\alpha}(t,X_t,\bar{\mu},\bar{\nu},Y_t)$. When solving the fixed point step, we identify $\bar{\nu}_t=\mathbb{E}(\hat{\alpha}_t)$. By taking the expectation, we find:
\begin{equation*}
    \bar{\nu}_t = \mathbb{E}(\hat{\alpha}_t)=c^{MFG}(t)\mathbb{E}(Y_t),
	\qquad 
	\text{with} 
	\quad
    c^{MFG}(t)=-\frac{b_2(t)}{r(t)+\bar{r}(t)(1-\bar{s}(t))}.
\end{equation*}
Thus, from equation (\ref{eq:MFG_opt_alpha_Pontryagin}) we must have:
\begin{equation}
    \hat{\alpha}_t=a^{MFG}(t) Y_t+b^{MFG}(t)\mathbb{E}(Y_t),
\label{eq:MFG_opt_alpha_Y_expY}
\end{equation}
with
\begin{equation*}
	a^{MFG}(t)=-\frac{b_2(t)}{r(t)+\bar{r}(t)}
	\qquad
	\text{and}
	\qquad
	b^{MFG}(t)=-\frac{\bar{r}(t)\bar{s}(t)b_2(t)}{(r(t)+\bar{r}(t))(r(t)+\bar{r}(t)(1-\bar{s}(t)))}.
\end{equation*}
Note that $c^{MFG}(t)=a^{MFG}(t)+b^{MFG}(t)$. The solution of the mean field game equilibrium problem is given by the solution to the FBSDE system:
\begin{equation}
\left\{
\begin{array}{lcl}
      \displaystyle  dX_t&=& \left[b_1(t)X_t+\bar{b}_1(t) \mathbb{E}X_t+a^{MFG}(t)b_2(t)Y_t+(b^{MFG}(t)b_2(t)+c^{MFG}(t)\bar{b}_2(t))\mathbb{E}Y_t\right]dt+\sigma dW_t \\[5pt]
        dY_t&=& -\left[(q(t)+\bar{q}(t))X_t-\bar{q}(t)s(t)\mathbb{E}X_t+b_1(t)Y_t \right]dt +Z_t dW_t,
\end{array}
\right.
\label{eq:FBSDE_EMFG}
\end{equation}
with initial condition $X_0=\xi$, a random variable with finite mean and variance, and terminal condition $Y_T=(q_T+\bar{q}_T)X_T-\bar{q}_Ts_T\mathbb{E}X_T$.

@@

It will be useful for us to note here the scalar Riccati equations associated with $u_t := \lambda \bar{\eta}_t^{MFG}$, $w_t := \bar{\eta}_t^{MKV}$ and $\eta_t$:
    \begin{IEEEeqnarray}{lLLLClClCl}
		\dot{u}_t &- 2A^u u_t &- B u_t^2 &+ C^u &= & 0, &\qquad& u_T &=& D^u,
		\label{eq:riccati_ut} \\
	    \dot{w}_t &- 2A^w w_t &- B w_t^2 &+ C^w &=&0, &\qquad &w_T &= & D^w,
		\label{eq:riccati_wt}\\
	    \dot{\eta}_t &- 2A^\eta \eta_t &- B^\eta \eta_t^2 &+ C^\eta &= & 0, &\qquad& \eta_T &= &D^\eta,
	\label{eq:riccati_etat}
	\end{IEEEeqnarray}
	with:
	\begin{IEEEeqnarray*}{rClcrClcrCl}
		A^u &=& - \left(b_1 + \frac{\bar{b}_1}{2}\right), & \qquad & 
		A^w &=& - (b_1 + \bar{b}_1), & \qquad & 
		A^\eta &=& - b_1, \nonumber \\
		B^u &=& \frac{(b_2+\bar{b}_2)^2}{r + \bar{r}(1-\bar{s})^2}, & \qquad &	
		B^w &=& \frac{(b_2+\bar{b}_2)^2}{r + \bar{r}(1-\bar{s})^2}, &\qquad & 
		B^\eta &=& \frac{b_2^2}{r+ \bar{r}}, \nonumber \\
		C^u &=& \lambda (q + \bar{q}(1-s)),& \qquad & 
		C^w &=& q + \bar{q}(1-s)^2, & \qquad &
		C^\eta &=& q + \bar{q}, \nonumber \\
		D^u &=& \lambda(q_T + \bar{q}_T(1-s_T)), & \qquad & 
		D^w &=& q_T + \bar{q}_T(1-s_T)^2, & \qquad & 
		D^\eta &=& q_T + \bar{q}_T .
	\end{IEEEeqnarray*}
If $B^u\neq 0$, $B^u D^u \geq 0$ and $B^uC^u >0$, we have (see equation (\ref{eq:riccati_ut_sol_app}) in Appendix A) the existence and uniqueness for $u_t$ which can be expressed by:
\begin{equation}
		u_t= \frac{C^u(1-e^{-(\delta_u^+ - \delta_u^-)(T-t)})+D^u(\delta^+_u -\delta^-_ue^{-(\delta_u^+ - \delta_u^-)(T-t)} )}{B^uD^u(1-e^{-(\delta_u^+ - \delta_u^-)(T-t)})+ \delta^+_ue^{-(\delta_u^+ - \delta_u^-)(T-t)} -\delta^-_u  },
	\label{eq:riccati_ut_sol}
\end{equation}
with $\delta^\pm_u = -A^u \pm \sqrt{(A^u)^2 + B^u C^u}$. Under assumption (\ref{eq:assumptions}), the above conditions on $B^u$, $C^u$, and $D^u$ are satisfied, and we have $\delta^-_u<0<\delta^+_u$, $u_t > 0$ for all $t\in[0,T)$, and $u_T \geq 0$. We have analogous expressions for $w_t$ and $\eta_t$, in terms of $\delta^\pm_w$ and $\delta^\pm_{\eta}$, respectively. Note that $B^u=B^w=:B$.

@@

We study in the following the variation of $PoA$ by letting only one of the coefficients tend to zero or to infinity. In order to make the following computations easier to follow, we repeat equations (\ref{eq:PoA_LQ}), (\ref{eq:x_bar_MFG_explicit}), and (\ref{eq:v_t}), which we recall is equivalent to  equation (\ref{eq:v_t_MKV}), using the above notations:
\begin{IEEEeqnarray}{l}
	PoA = 1+ \frac{\Delta SC}{SC^{MKV}}= 1 +  \frac{ \displaystyle B\int_0^T (u_t-w_t)^2 \cdot (\bar{x}_t^{MFG})^2 dt }{\displaystyle \int_0^T \left[q+\bar{q} + B^{\eta} \eta_t^2 \right] v_t dt + (q_T + \bar{q}_T) v_T + w_0 (\mathbb{E}(\xi))^2},
\label{eq:PoA_LQ_new_notation} \\[3pt]
	\displaystyle \bar{x}^{MFG}_t =\mathbb{E}(\xi) e^{\int_0^t(b_1+\bar{b}_1-Bu_s)ds},
\label{eq:x_bar_MFG_explicit_new_notation} \\[3pt]	
	v_t = Var(\xi)e^{\int_0^t 2(b_1-B^{\eta} \eta_s)ds}+\sigma^2 \int_0^t e^{2 \int_s^t (b_1-B^{\eta}\eta_u) du}ds.
\label{eq:v_t_new_notation}
\end{IEEEeqnarray}
Also for convenience, recall the definition from equation (\ref{eq:lambda}):
$
\displaystyle
\lambda = \frac{b_2}{b_2 + \bar{b}_2}\cdot \frac{ r + \bar{r}(1-\bar{s})^2 }{r + \bar{r}(1-\bar{s})}.
$

\noindent In the following propositions, we utilize the following assumption to make their proofs simpler.	
\begin{assumption}\label{assumption}
	Assume (\ref{eq:assumptions}). In addition, assume: $b_1 > 0$, $D^{u} > 0$, $D^w > 0$, $D^{\eta} >0$ and the initial condition satisfies $\mathbb{E}(\xi) \neq 0$.
\end{assumption}
